---
title: "EDA"
author: "Serkan Oral"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, message=TRUE, warning=TRUE, include=FALSE}
# Packages

library(tidyverse)
library(data.table)
library(naniar)
library(janitor)
library(patchwork)

```

We will mostly focus on the platform people learn data science.
We may check other features on some other notebook.

# Data Cleaning 

Downloading the data.
```{r}
kaggle <- fread("kaggle_survey_2022_responses.csv", skip = 1)
```
We add surveyer_id column, it will be usefull.

```{r}
kaggle$surveyer_id <- seq.int(nrow(kaggle))

kaggle <- kaggle %>% 
  select(surveyer_id,everything()) 
```

# Platforms

We extract just some columns, where we do our analysis.

```{r}

platform_data <- kaggle[, 1:18 ]


```


We make the dataset to longer format, cause surveyers can choose more than one platform.

```{r}
platform_data_melt <-  melt(platform_data,  measure.vars = grep("^On which platforms", colnames(kaggle))) 


```


Changing NA'S to "No Response"

```{r}
platform <- platform_data_melt %>% 
  select(-variable) %>% 
  rename(Platform_name = "value") %>% 
  mutate(Platform_name = ifelse(Platform_name == "","No Response", Platform_name)) %>% 
  mutate_if(is.character, factor) %>% 
  clean_names()
```

Here on there will some plots for you understand the relationship.

```{r}

platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  ggplot(aes(fct_infreq(platform_name), fill = what_is_your_age_number_years) )+
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab(NULL) + ylab(NULL) + labs(fill = "Age")

```

There is no student in this plot, I want to see that if this change anythink.

```{r}
platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  filter(are_you_currently_a_student_high_school_university_or_graduate == "No") %>% 
  ggplot(aes(fct_infreq(platform_name), fill = what_is_your_age_number_years) )+
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab(NULL) + ylab(NULL) + labs(fill = "Age")
```


```{r}

platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  ggplot(aes(fct_infreq(platform_name), fill = are_you_currently_a_student_high_school_university_or_graduate) )+
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab(NULL) + ylab(NULL) + labs(fill = "In School or University")
```



We get the Country name big to small according to number of replay to country.

```{r}

country_names <- platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by(in_which_country_do_you_currently_reside) %>% summarise(n = n()) %>% arrange(desc(n)) %>%  pull(in_which_country_do_you_currently_reside)

```

First 3 Countries
```{r}
country_names[1:3]

platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  filter(in_which_country_do_you_currently_reside %in% country_names[1:3]) %>% 
  ggplot(aes(fct_rev(fct_infreq(in_which_country_do_you_currently_reside)), fill = fct_infreq(platform_name)) )+
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab(NULL) + ylab(NULL) + labs(fill = "Age") +coord_flip()
```

Later 10 Countries
```{r}
country_names[1:3]

platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  filter(in_which_country_do_you_currently_reside %in% country_names[4:14]) %>% 
  ggplot(aes(fct_rev(fct_infreq(in_which_country_do_you_currently_reside)), fill = fct_infreq(platform_name)) )+
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab(NULL) + ylab(NULL) + labs(fill = "Age") +coord_flip()
```

Later 10 Countries
```{r}
country_names[1:3]

platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  filter(in_which_country_do_you_currently_reside %in% country_names[15:25]) %>% 
  ggplot(aes(fct_rev(fct_infreq(in_which_country_do_you_currently_reside)), fill = fct_infreq(platform_name)) )+
  geom_bar() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab(NULL) + ylab(NULL) + labs(fill = "Age") +coord_flip()
```



How many percentage each platform gets from the market.

```{r}
platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by(platform_name) %>% 
  dplyr::summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3)) %>% 
  arrange(desc(freq)) %>% 
  mutate(platform_name = factor(platform_name)) %>% 
  ggplot(aes(fct_rev(fct_reorder(platform_name , freq)), freq, fill = fct_rev(fct_reorder(platform_name , freq)))) +
  geom_col() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab(NULL)+ ylab("Percentage") + labs(fill = "Platform Name") +
  scale_y_continuous(labels = scales::percent)
```

Where kaggle courses are number 1

```{r}
platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by( in_which_country_do_you_currently_reside,platform_name) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3), rank = rank(desc(freq), ties.method = "first")) %>% 
  arrange(in_which_country_do_you_currently_reside, desc(freq)) %>%  
  filter(platform_name == "Kaggle Learn Courses" & rank == 1) %>% 
  ggplot(aes(fct_rev(fct_reorder(in_which_country_do_you_currently_reside,freq )), freq,
             fill = fct_rev(fct_reorder(in_which_country_do_you_currently_reside,freq )))) +
  geom_col() + xlab("Percentage") + ylab(NULL) + labs(fill = "Countries - Kaggle is their first platform") +scale_y_continuous(labels = scales::percent)
```

We will extract where kaggle is in the first five place.

```{r}
kaggle_first_countries <- platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by( in_which_country_do_you_currently_reside,platform_name) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3), rank = rank(desc(freq), ties.method = "first")) %>% 
  arrange(in_which_country_do_you_currently_reside, desc(freq)) %>%  
  filter(platform_name == "Kaggle Learn Courses" & rank == 1) %>% 
  pull(in_which_country_do_you_currently_reside)

kaggle_second_countries <- platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by( in_which_country_do_you_currently_reside,platform_name) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3), rank = rank(desc(freq), ties.method = "first")) %>% 
  arrange(in_which_country_do_you_currently_reside, desc(freq)) %>%  
  filter(platform_name == "Kaggle Learn Courses" & rank == 2) %>% 
  pull(in_which_country_do_you_currently_reside)


kaggle_third_countries <- platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by( in_which_country_do_you_currently_reside,platform_name) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3), rank = rank(desc(freq), ties.method = "first")) %>% 
  arrange(in_which_country_do_you_currently_reside, desc(freq)) %>%  
  filter(platform_name == "Kaggle Learn Courses" & rank == 3) %>% 
  pull(in_which_country_do_you_currently_reside)

kaggle_fourth_countries <- platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by( in_which_country_do_you_currently_reside,platform_name) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3), rank = rank(desc(freq), ties.method = "first")) %>% 
  arrange(in_which_country_do_you_currently_reside, desc(freq)) %>%  
  filter(platform_name == "Kaggle Learn Courses" & rank == 4) %>% 
  pull(in_which_country_do_you_currently_reside)


kaggle_fifth_countries <- platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  group_by( in_which_country_do_you_currently_reside,platform_name) %>% 
  summarise(cnt = n()) %>% 
  mutate(freq = round(cnt/sum(cnt),3), rank = rank(desc(freq), ties.method = "first")) %>% 
  arrange(in_which_country_do_you_currently_reside, desc(freq)) %>%  
  filter(platform_name == "Kaggle Learn Courses" & rank == 5) %>% 
  pull(in_which_country_do_you_currently_reside)

```
Countries where kaggle has the first place

```{r}
platform %>% 
  filter(platform_name != "No Response" &platform_name != "Other" 
         & platform_name != "None") %>% 
  filter(in_which_country_do_you_currently_reside %in% kaggle_first_countries) %>% 
  group_by(in_which_country_do_you_currently_reside) %>% 
  summarise(cnt = n(),platform_name) %>% 
  mutate(freq = round(cnt/sum(cnt),3)) %>% 
  arrange(desc(freq)) %>% 
  ggplot(aes(in_which_country_do_you_currently_reside,freq, fill = fct_infreq(platform_name) )) + geom_col(position = "fill") + xlab(NULL) + ylab("Percentage") + labs(fill = "Platforms")  +scale_y_continuous(labels = scales::percent)
```


# Conclusion

Kaggle learn courses are doing the something right on above countries. They may have apply similar strategy to increase their sales where they are on below first. 


